---
title: "Tidyverse Extend"
author: "Molly Siebecker"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(readr)
spotify_data <- read_csv(url('https://raw.githubusercontent.com/acatlin/FALL2023TIDYVERSE/main/spotify_songs.csv'))
```

# Initial Code

The code below was provided by Fomba Kassoh. He demonstrated multiple examples of using ggplot2. I will extend only one of his examples, that of creating a box plot, to show how additional ggplot2 aesthetic features may be applied.

```{r}
library(ggplot2)

# Assuming your data is in a data frame named 'df'
p <- ggplot(spotify_data, aes(x = playlist_genre, y = track_popularity)) + 
  geom_boxplot() +
  theme_minimal() + 
  labs(title = "Distribution of Track Popularity by Playlist Genre",
       x = "Playlist Genre",
       y = "Track Popularity")

print(p)
```

# Extension

Below I will recreate Fomba's boxplot and add additional layers to the code.

We could opt to change the coordinate system, in this case, flipping the x- and y-axes, using coord_flip().
```{r}
## coordinate system
ggplot(spotify_data, aes(x = playlist_genre, y = track_popularity)) + 
  geom_boxplot() +
  theme_minimal() + 
  coord_flip() +
  labs(title = "Distribution of Track Popularity by Playlist Genre",
       x = "Playlist Genre",
       y = "Track Popularity")
```
 
We can also select from pre-set color palettes, using "fill" and "scale_fill_brewer."

```{r}

## view available color palettes

RColorBrewer::display.brewer.all()

## add color palette
ggplot(spotify_data, aes(x = playlist_genre, y = track_popularity, fill = playlist_genre)) + 
  geom_boxplot() +
  theme_minimal() + 
  coord_flip() +
  scale_fill_brewer(palette = "YlOrRd")+
  labs(title = "Distribution of Track Popularity by Playlist Genre",
       x = "Playlist Genre",
       y = "Track Popularity")
```

We could even change the background using theme, which was previously set to theme_minimal.

```{r}
ggplot(spotify_data, aes(x = playlist_genre, y = track_popularity, fill = playlist_genre)) + 
  geom_boxplot() +
  theme_dark() + 
  coord_flip() +
  scale_fill_brewer(palette = "YlOrRd")+
  labs(title = "Distribution of Track Popularity by Playlist Genre",
       x = "Playlist Genre",
       y = "Track Popularity")
```

We could also remove the legend using "guides," since it is not adding any new information.

```{r}
ggplot(spotify_data, aes(x = playlist_genre, y = track_popularity, fill = playlist_genre)) + 
  geom_boxplot() +
  theme_dark() + 
  coord_flip() +
  guides(fill = "none") +
  scale_fill_brewer(palette = "YlOrRd")+
  labs(title = "Distribution of Track Popularity by Playlist Genre",
       x = "Playlist Genre",
       y = "Track Popularity")
```

Finally, in addition to the title and x- and y-axis labels, we can add a subtitle and caption.

```{r}
ggplot(spotify_data, aes(x = playlist_genre, y = track_popularity, fill = playlist_genre)) + 
  geom_boxplot() +
  theme_dark() + 
  coord_flip() +
  guides(fill = "none") +
  scale_fill_brewer(palette = "YlOrRd")+
  labs(title = "Most Popular Genres",
       x = "Playlist Genre",
       y = "Track Popularity",
       subtitle = "Distribution of Track Popularity by Playlist Genre",
       caption = "Pop, Latin songs achieve greatest median popularity.")
```

# Conclusion

The ability to add layers in ggplot2 provides numerous opportunities for customization. There are additional aesthetic arguments not demonstrated here, as they would not be appropriate for this example.  
